---
title: "753 problem Set 6"
author: "Jesús Lara"
date: "12/7/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Prepare Dataset, include=FALSE}
#Some packages

library(dplyr)
library(tidyr)
library(data.table)
library(ggplot2)
library("rio")
library(matlib)
library(gdata)
library(scales)
library(ggplot2)
library(foreign)
library(rmarkdown)
library(fastDummies)
library(stargazer)
library(naniar)


options(scipen=10000)
options(digits=4)

rm(list=ls())


#Set my working directory
setwd("C:/Users/User/Documents/GitHub/Problem-Sets--753/PS6")

#Import data on employment 2005-2015
emp.2015<-import("nal2005.xls")



#Import data on employment 1995-2004
#Route INEGI 2004
# https://www.inegi.org.mx/sistemas/bie/?idserPadre=10100035
emp.2004<-import("nal.xls")
#Change colnames
colnames(emp.2004)<-emp.2004[3,]

#Keep relevant variables 1
emp.2004<-data.frame(emp.2004[-c(22:138),])

#Keep relevant variables2:
emp.2004<-emp.2004 %>% filter(Indicador=="Población económicamente activa (PEA)"|
                                Indicador=="Trabajadores por cuenta propia" |
                                Indicador=="Trabajadores no remunerados"|
                                Indicador=="Tasa de condiciones críticas de ocupación (TCCO)"|
                                Indicador=="Tasa de trabajo asalariado"|
                                Indicador=="Tasa de presión general (TPRG)"|
                                Indicador=="Tasa de ocupación parcial y desocupación 1 (TOPD1)"|
                                Indicador=="Tasa de desocupación"|
                                Indicador=="Tasa de participación"|
                                Indicador=="Cuenta propia")
#Change colnames
years1<-c(1995:1999)
#2000 only 3 quartes
year2000<-paste(2000,c(2:4), sep=".")
years2<-c(2001:2004)
quarters<-c(1:4)
years2.final<-paste(rep(years2,each=4), quarters, sep=".")
years<-c(years1,year2000,years2.final)
years
colnames(emp.2004)<-c("indicador", years)

#Convert to numeric
emp.2004<-emp.2004 %>% mutate(across(2:25,as.numeric))

#Take average of years for which we have quarterly data
emp.2004.2<-emp.2004[c(1,7:25)]

emp.2004.2<-emp.2004.2 %>% pivot_longer(cols=c(2:20),names_to="year")

#Split year and quarter
emp.2004.2<- emp.2004.2 %>% 
  separate(year,c("year","quarter",sep="."))

#Delete "." columns
emp.2004.2<-emp.2004.2 %>% select(-".")

#Take average
emp.2004.2<-emp.2004.2 %>% group_by(indicador,year) %>% 
  summarise(across(value, mean, na.rm=TRUE)) %>% ungroup()

#Pivot_longer first part
emp.2004.1<-emp.2004 %>% select(c(1:6))
emp.2004.1<-emp.2004.1 %>% pivot_longer(cols=(2:6), names_to="year")

#Append both objects
emp.2004<-bind_rows(emp.2004.1,emp.2004.2)

#Now pivot wider: we want our variables as columns
emp.2004<-emp.2004 %>% pivot_wider(names_from = indicador)

#Rename cols (Also can use rename)
colnames(emp.2004)<-c("year",
                      "pea",
                      "cuenta.propia.total",
                      "no.remunerados.total",
                      "participacion",
                      "desocupacion",
                      "topd1",
                      "tprg",
                      "asalariado",
                      "tcco")

#Calculate tasas
emp.2004 <- emp.2004 %>% 
  mutate(cuenta.propia=(cuenta.propia.total/pea)*100,
         no.remunerados=(no.remunerados.total/pea)*100)


emp.2004<-emp.2004 %>% select(-pea,-cuenta.propia.total,-no.remunerados.total)
################################################
#Now these same variables but for 2005-2015#####

#Change colnames
colnames(emp.2015)<-emp.2015[2,]

#Delete rows 1 and 2
emp.2015<-emp.2015 %>% 
  slice(-c(1,2))

#Delete last rows
emp.2015<-emp.2015 %>% 
  slice(-c(191:195))

#Separate year month
emp.2015<- emp.2015 %>% 
  separate(Periodo,c("year","month"),sep="/")

#Make numeric
emp.2015<-emp.2015 %>% 
  mutate(across(3:14,as.numeric))

#Take year average
emp.2015<-emp.2015 %>% group_by(year) %>% 
  summarise(across(where(is.numeric), mean)) %>% ungroup()

#Rename renumerados lol
emp.2015<- emp.2015 %>% rename(no.remunerados=no.renumerados)
######### Append to have all employment data ######
emp<-bind_rows(emp.2004,emp.2015)


########################################################
#Import data Labor Income Share
LIS<-import("LIS.xlsx")
#### Clean data on the LIS

#First keep only data on WS and LIS
LIS.w<-LIS[,1:19]

#Get vector of colnames
names<-c("WS","LIS1","LIS2")
cats<-c("WE","PBS","Tradables","Non-tradables","Agriculture", "Manufacturing")
names.cats<-paste(names,rep(cats,each=3),sep=".")
names.cats
colnames(LIS.w)<-c("year",names.cats)

#Remove 1st 2 rows
LIS.w<-LIS.w %>% slice(-c(1,2))

#Make everything numeric
LIS.w<- LIS.w %>% mutate(across(where(is.character),as.numeric))

#Merge with data on employmeny
emp.LIS<-merge(emp, LIS.w, by="year",all=TRUE)



###############################
#### Import inflation data ####
###############################
inflation<-import("inflation.xlsx")
colnames(inflation)<-c("year","infl.IPC","infl.suby","infl.no.suby")

#separate year
inflation<-inflation %>%separate(year,c("year","month"),"-") 

#get year averages
inflation<-inflation %>% group_by(year) %>% summarise(across(starts_with("infl"),mean)) %>% ungroup()

#Merge with all our previous data
complete<-merge(emp.LIS,inflation, by="year",all=TRUE)
complete<-complete %>% arrange(year)

#Make labels
years.5<-c("1990",NA,NA,NA,NA,"1995",NA,NA,NA,NA,"2000",NA,NA,NA,NA,"2005",NA,NA,NA,NA,"2010",NA,NA,NA,NA,"2015",NA,NA,NA,NA,"2020")
complete$years.5<-years.5
```

# Inflation, Unemployment and Class Struggle in Neoliberal Mexico

In this report I will explore the relationship between inflation, unemployment and distribution of value added in Mexico over the last 10-15 years. The motivation comes from Ibarra & Ros (2018) study about the decline of the Labor Income Share in Mexico during the period 1990-2015, a pattern that contrasts with what happened in most Latin American countries during that period. Their theoretical analysis is based on structuralist models of economic development, that start with the structural heterogenity that characterizes Third World Countries by recongnizing the existence of at least two different economic sectors: the modern, capitalist one and a "traditional" or subsistence sector. 

```{r message=FALSE, warning=FALSE}
f1<-complete %>% arrange(year) %>% ggplot(aes(x=desocupacion, 
             y=infl.IPC, label=years.5))
f1<-f1+ geom_point() 
f1<-f1+ geom_text(data=complete %>% filter(!is.na(years.5)),aes(label=year),
                  angle=45,hjust=-0.25,vjust=-0.25)
f1<-f1+ theme_bw()
f1<-f1+ geom_path()
f1<-f1+scale_x_continuous(name="Unemployment Rate (%)", 
                        breaks=c(2:8),limits=c(2,7.5))
f1<-f1+scale_y_continuous(name="Inflation (CPI) (%)",limits=c(2,40))
f1<-f1+ ggtitle("Inflation and Unemployment Rate: 1995-2020")

plot(f1)
```

```{r, Inflation and TOPD1, warning=FALSE}

# 2nd: TCCP
f1<-complete %>% arrange(year) %>% ggplot(aes(x=topd1, 
                                              y=infl.IPC, label=years.5))
f1<-f1+ geom_point() 
f1<-f1+ geom_text(data=complete %>% filter(!is.na(years.5)),aes(label=year),
                  angle=45,hjust=-0.25,vjust=-0.25)
f1<-f1+ theme_bw()
f1<-f1+ geom_path()
f1<-f1+scale_x_continuous(name="Rate of Unemployment and Partial Ocupation (%)", 
                          )
f1<-f1+scale_y_continuous(name="Inflation (CPI) (%)",limits=c(0,40))
f1<-f1+ ggtitle("Inflation and RUPO: 1995-2020")
plot(f1)

```

```{r, Wage Share and Unemployment, warning=FALSE}
f1<-complete %>% arrange(year) %>% ggplot(aes(x=desocupacion, 
                                              y=WS.WE))
f1<-f1+ geom_point() 
f1<-f1+ geom_text(data=complete %>% filter(!is.na(years.5)),aes(label=year),
                  angle=90,hjust=-0.25,vjust=-0.25)
f1<-f1+ theme_bw()
f1
f1<-f1+scale_x_continuous(name="Unemployment rate")
f1<-f1+scale_y_continuous(name="Wage Share", limits=c(25,31))
f1<-f1+ ggtitle("Unemployment and Wage Share in Mexico: 1995-2015")
f1<-f1+geom_path()
plot(f1)
```

```{r, Wage Share and TOPD1, warning=FALSE}
f1<-complete %>% arrange(year) %>% ggplot(aes(x=topd1, 
                                              y=WS.WE))
f1<-f1+ geom_point() 
f1<-f1+ geom_text(data=complete %>% filter(!is.na(years.5)),aes(label=year),
                  angle=90,hjust=-0.25,vjust=-0.25)
f1<-f1+ theme_bw()

f1<-f1+scale_x_continuous(name="Partial employment and unemploymante rate")
f1<-f1+scale_y_continuous(name="Wage Share", limits=c(25,31))
f1<-f1+ ggtitle("Partial employment and unemployment vs Wage Share in Mexico: 1995-2015")
f1<-f1+geom_path()
f1<-f1+ggtitle("Wage Share vs Partial employment and unemployment rate")
plot(f1)
```

```{r,Wage Share manufacturing and TOPD1, warning=FALSE}

f1<-complete %>% arrange(year) %>% ggplot(aes(x=topd1, 
                                              y=WS.Manufacturing))
f1<-f1+ geom_point() 
f1<-f1+ geom_text(data=complete %>% filter(!is.na(years.5)),aes(label=years.5),
                  angle=90,hjust=-0.25,vjust=-0.25)
f1<-f1+ theme_bw()
f1
f1<-f1+scale_x_continuous(name="Partial Employment and Unemploymante Rate")
f1<-f1+scale_y_continuous(name="Wage Share", limits=c(17,29))
f1<-f1+ ggtitle("PEUR vs Wage Share in Manufacturing: Mexico 1995-2015")
f1<-f1+geom_path()
plot(f1)
```

```{r, Wage Share manufacturing and TPRG, warning=FALSE}

f1<-complete %>% arrange(year) %>% ggplot(aes(x=tprg, 
                                              y=WS.Manufacturing))
f1<-f1+ geom_point() 
f1<-f1+ geom_text(data=complete %>% filter(!is.na(years.5)),aes(label=years.5),
                  angle=90,hjust=-0.25,vjust=-0.25)
f1<-f1+ theme_bw()
f1
f1<-f1+scale_x_continuous(name="General Pressure Rate")
f1<-f1+scale_y_continuous(name="Wage Share", limits=c(17,29))
f1<-f1+ ggtitle("General Pressure Rate vs Wage Share in Manufacturing: Mexico 1995-2015")
f1<-f1+geom_path()
plot(f1)



```

```{r, Wage share tradables vs self-employed, , warning=FALSE} 
f1<-complete %>% arrange(year) %>% ggplot(aes(x=cuenta.propia, 
                                              y=WS.Tradables))
f1<-f1+ geom_point() 
f1<-f1+ geom_text(data=complete %>% filter(!is.na(years.5)),aes(label=years.5),
                  angle=90,hjust=-0.25,vjust=-0.25)
f1<-f1+ theme_bw()
plot(f1)
f1<-f1+scale_x_continuous(name="Self-employed",limits=c(21,25))
f1<-f1+scale_y_continuous(name="Wage Share in the Tradable sector", limits=c(17,27))
f1<-f1+ ggtitle("Self-employment vs Wage Share in Manufacturing: Mexico 1995-2015")
f1<-f1+geom_path()
plot(f1)



``` 